\hl{Identifying Effective Motivational Interviewing Communication Sequences Using Automated Pattern Analysis}

### 2018-06-28 jhir/latex-manuscript/manuscript.tex
prepared final copy for the minor revision

\subsection{Dataset}
\label{subsec:dataset}
The \hl{dataset} for this study consists of 37 audio recorded transcripts from Motivational Interviewing (MI) sessions conducted with African American adolescents seeking weight loss. Each transcript represented a one-on-one interaction between the adolescent and the MI counselor followed by a one-on-one interaction between the caregiver and the MI counselor. These encounters \hl{had been previously} annotated using the Minority Youth Sequential Code for Observing Process Exchanges (MYSCOPE)~\cite{carcone2013provider} \hl{excluding conversations correspond to greetings, farewell and interview setups such as table and camera settings.} The MYSCOPE is an adaptation of the original MI-SCOPE~\cite{martin2005motivational}, a qualitative code scheme to characterize patient-counselor communication during MI treatment sessions. The MYSCOPE was informed by MI fidelity code schemes including the MI Treatment Integrity Scale (MITI)~\cite{moyers2005assessing}, the MI Skill Code (MISC)~\cite{apodaca2007motivational} and Amrhein's conceptualization of change talk and commitment~\cite{amrhein2004does}. A primary coder used the MYSCOPE to annotate transcripts of all 37 sessions; a secondary coder coded 20\% ($n = 7$) of the sessions for inter-rater reliability which was good ($k = .696$). The experimental dataset consists of 7,192 patient, caregiver and counselor utterances segmented and annotated with the MYSCOPE behavior codes, illustrated in Table~\ref{tab:codebook}.

@@

\subsubsection{Frequent Pattern Mining}
\label{subsubsec:frequentPatternMining}
We applied frequent pattern mining to identify frequently occurring \hl{patterns of patient-counselor} behavior codes in successful and unsuccessful communication sequences. Behavior codes in these patterns may be separated by one or more other codes. For this purpose, we utilized FPClose~\cite{grahne2005fast}, an efficient state-of-the-art \hl{closed} frequent pattern mining algorithm implemented in SPMF~\cite{fournier2014spmf, fournier2016spmf}, to identify frequent patterns of patient-counselor communication behaviors in successful and unsuccessful MI communication sequences. SPMF is an open-source library providing more than 150 data mining algorithms. Popular \hl{non-closed} frequent pattern mining algorithms include Apriori~\cite{agrawal1994fast} and FP-Growth~\cite{han2000mining}. A \textit{frequent pattern} is defined as a \hl{pattern of observations}, which appears in a given set of sequences more often than a user-specified threshold called the \textit{minimum support count}. For example, \{A\}, \{C\}, \{D\} and \{C, D\} are frequent patterns in the example set of sequences in Figure~\ref{fig:fpm}, since these patterns appear at least 2 times, which is the minimum support count in this example. In this work, we identified and analyzed \textit{closed frequent patterns} among all sequences of behavior codes in successful and unsuccessful communication sequences. A \textit{frequent pattern} is \textit{closed} if none of its supersets have the same support count~\cite{pasquier1999discovering}, where a set X is a superset of another set Y, if X contains all the elements of the set Y. For example, the itemsets \{A\}, \{C\}, \{D\} and \{C, D\} in Figure~\ref{fig:fpm} are not closed frequent \hl{patterns since} their supersets \{A, B\}, \{B, C\}, \{B, D\} and \{B, C, D\} have the same support count. Therefore, \{B\}, \{A, B\}, \{A, D\}, \{B, C\}, \{B, D\} and \{B, C, D\} are closed frequent \hl{patterns since} none of their supersets have the same support count. On the other hand, itemsets \{A, C\}, \{A, B, C\}, \{A, B, D\}, \{A, C, D\} and \{A, B, C, D\} have support counts of 1, 0, 1, 0 and 1, respectively, which is less than the minimum support count and thus are identified as non-frequent itemsets. Since the threshold for minimum support count depends on a task and is typically determined by the domain expert, we followed \hl{prior work}~\cite{moradi2017quantifying, liu1999mining} and set the minimum support count as 10\% of the total number of all communication sequences, which is 110 for successful and 25 for unsuccessful communication sequences. For each pattern, the statistical significance \hl{of the difference} between pattern occurrence in successful and unsuccessful communications was reported by computing \hl{p-value of Pearson's} chi-square test.

>>>>

\subsection{Dataset}
\label{subsec:dataset}
The \hl{dataset} for this study consists of 37 audio recorded transcripts from Motivational Interviewing (MI) sessions conducted with African American adolescents seeking weight loss. Each transcript represented a one-on-one interaction between the adolescent and the MI counselor followed by a one-on-one interaction between the caregiver and the MI counselor. These encounters \hl{had been previously} annotated using the Minority Youth Sequential Code for Observing Process Exchanges (MYSCOPE)~\cite{carcone2013provider} \hl{excluding conversations that correspond to greetings, farewell and interview setups such as table and camera settings.} The MYSCOPE is an adaptation of the original MI-SCOPE~\cite{martin2005motivational}, a qualitative code scheme to characterize patient-counselor communication during MI treatment sessions. The MYSCOPE was informed by MI fidelity code schemes including the MI Treatment Integrity Scale (MITI)~\cite{moyers2005assessing}, the MI Skill Code (MISC)~\cite{apodaca2007motivational} and Amrhein's conceptualization of change talk and commitment~\cite{amrhein2004does}. A primary coder used the MYSCOPE to annotate transcripts of all 37 sessions; a secondary coder coded 20\% ($n = 7$) of the sessions for inter-rater reliability which was good ($k = .696$). The experimental dataset consists of 7,192 patient, caregiver and counselor utterances segmented and annotated with the MYSCOPE behavior codes, illustrated in Table~\ref{tab:codebook}.

@@

\subsubsection{Frequent Pattern Mining}
\label{subsubsec:frequentPatternMining}
We applied frequent pattern mining to identify frequently occurring \hl{patterns of patient-counselor} behavior codes in successful and unsuccessful communication sequences. Behavior codes in these patterns may be separated by one or more other codes. For this purpose, we utilized FPClose~\cite{grahne2005fast}, an efficient state-of-the-art \hl{closed} frequent pattern mining algorithm implemented in SPMF~\cite{fournier2014spmf, fournier2016spmf}, to identify frequent patterns of patient-counselor communication behaviors in successful and unsuccessful MI communication sequences. SPMF is an open-source library providing more than 150 data mining algorithms. Popular \hl{non-closed} frequent pattern mining algorithms include Apriori~\cite{agrawal1994fast} and FP-Growth~\cite{han2000mining}. A \textit{frequent pattern} is defined as a \hl{pattern of observations}, which appears in a given set of sequences more often than a user-specified threshold called the \hl{\textit{minimum support count}}. For example, \{A\}, \{C\}, \{D\} and \{C, D\} are frequent patterns in the example set of sequences in Figure~\ref{fig:fpm}, since these patterns appear at least 2 times, which is the minimum support count in this example. In this work, we identified and analyzed \textit{closed frequent patterns} among all sequences of behavior codes in successful and unsuccessful communication sequences. A \textit{frequent pattern} is \textit{closed} if none of its supersets have the same support count~\cite{pasquier1999discovering}, where a set X is a superset of another set Y, if X contains all the elements of the set Y. For example, the itemsets \{A\}, \{C\}, \{D\} and \{C, D\} in Figure~\ref{fig:fpm} are not closed frequent \hl{patterns since} their supersets \{A, B\}, \{B, C\}, \{B, D\} and \{B, C, D\} have the same support count. Therefore, \{B\}, \{A, B\}, \{A, D\}, \{B, C\}, \{B, D\} and \{B, C, D\} are closed frequent \hl{patterns since} none of their supersets have the same support count. On the other hand, itemsets \{A, C\}, \{A, B, C\}, \{A, B, D\}, \{A, C, D\} and \{A, B, C, D\} have support counts of 1, 0, 1, 0 and 1, respectively, which is less than the minimum support count and thus are identified as non-frequent itemsets. Since the threshold for minimum support count depends on a task and is typically determined by the domain expert, we followed \hl{prior work}~\cite{moradi2017quantifying, liu1999mining} and set the minimum support count as 10\% of the total number of all communication sequences, which is 110 for successful and 25 for unsuccessful communication sequences. For each pattern, the statistical significance \hl{of the difference} between pattern occurrence in successful and unsuccessful communications was reported by computing \hl{p-value of Pearson's} chi-square test.
